---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cyclinbayes

<!-- badges: start -->
<!-- badges: end -->

Cyclinbayes is an R package implementing bayesian methods for estimating both directed acyclic graphs (DAGs) and directed cyclic graphs (DCGs). The package provides full posterior inference for graph structures and causal effects using a hierarchical Bayesian model, allowing principled uncertainty quantification, edge inclusion probabilities, and credible intervals. 

For DAGs, cyclinbayes uses a hybrid MCMC scheme that combines collapsed Gibbs sampling with simulated annealing to improve mixing and avoid local optima. For DCGs, the package performs joint updates of adjacency and causal effect coefficients using random walk proposals, enabling inference in systems with feedback cycles. For DCGs, the package performs joint updates of adjacency and causal effect coefficients using random walk proposals, enabling inference in systems with feedback cycles. In both settings, sparsity is effectively recovered using spike and slab priors. 

Implemented in Rcpp, cyclinbayes leverages optimized C++ routines to handle large scale, high dimensional datasets. 

## Installation from Github
You can install the development version of cyclinbayes from GitHub using remotes:

```{r}
install.packages("remotes")
remotes::install_github("roblee01/cyclinbayesrev")
```

Then load the package:

```{r}
library(cyclinbayesrev)
library(ggplot2)
library(igraph)

# If using dist_type = "sid" and you see missing RBGL/graph:
#install.packages("BiocManager")
#BiocManager::install(c("graph", "RBGL"))
```

## Acyclic (DAG) Example

Below is a simple example demonstrating how to use the Bayesian LiNGAM
(DAG) sampler. Let \(p\) denote the number of variables, \(N\) the sample
size, and `num_iter` the number of MCMC iterations. Let
\(i \in \{1,\ldots,p\}\) index variables and
\(q \in \{1,\ldots,N\}\) index observations.

We generate structural errors from a finite Gaussian mixture model,

\[
\epsilon_i^{(q)}
\sim
\sum_{k=1}^{M}
\pi_{ik}\,N(\mu_{ik},\tau_{ik}),
\]

with

\[
M=2,\qquad
(\mu_{i1},\mu_{i2})=(-0.5,0.5),\qquad
(\tau_{i1},\tau_{i2})=(0.1,0.3),\qquad
(\pi_{i1},\pi_{i2})=(0.5,0.5).
\]

We generate a sparse DAG by including each candidate directed edge with
probability \(1-\Delta=0.1\), where \(\Delta=0.9\), while enforcing
acyclicity. For each included edge \(j\to i\), the corresponding nonzero
causal-effect coefficient is generated according to

\[
|B_{ij}| \sim \mathrm{Unif}(0.4,0.9),
\]

with its sign chosen independently to be positive or negative with equal
probability. For excluded edges, \(B_{ij}=0\).

Given the causal-effect matrix \(B\) and structural error matrix
\(\epsilon\), the observed data are generated from

\[
Y=(I-B)^{-1}\epsilon,
\]

where the \(i\)-th row of \(Y\) corresponds to

\[
(Y_i^{(1)},\ldots,Y_i^{(N)})^\top.
\]

```{r}
#######################################
# Simulation and MCMC settings
#######################################
N = 200 # Sample size for the test data
num_covariates = 10 # Number of features for test data
M = 5 # Number of finite clusters for mixed normal in likelihood
num_iter = 80000 # Number of iterations MCMC runs

#######################################
# Hyperparameter setup
#######################################
params = list(
  a_mu      = 0,
  b_mu      = 2,
  a_gamma   = 0.5,
  b_gamma   = 0.5,
  a_gamma_1 = 2,
  b_gamma_1 = 1,
  a_og_tao = 0.01,
  b_og_tao = 0.01,
  a_tao     = 2,
  b_tao     = 1,
  alpha     = 1
)

#######################################
# Generate DAG Example
#######################################
example_list = generates_examples_DAG(
  num_covariates = num_covariates,
  N              = N,
  M_input        = 2,        # components used to GENERATE the errors
  prob_sparsity  = 0.90,     # = edge_prob 0.10
  seed_input     = 21, # the script's set.seed(20 + sim)
  mag_range      = c(0.4, 0.9),
  prob_positive  = 0.5,
  seed_structure = 1,        # dag_structure(seed = 1)
  seed_weights   = 20        # dag_weights(seed = 20)
)

data_matrix = example_list$data_matrix
Adjacency_matrix_true = example_list$Adjacency_matrix_true
Causal_effect_matrix_true = example_list$Causal_effect_matrix_true
Z_matrix_true = example_list$Z_matrix_true
```

Before examining posterior summaries, it is helpful to visualize the true underlying DAG used in the simulation. This provides a direct point of comparison for the estimated graph structures returned by the sampler. The plot below displays the ground truth adjacency structure, where each directed edge represents a causal relationship from one variable to another.

```{r}
#######################################
# Build directed graph from adjacency matrix
#######################################

g_true = igraph::graph_from_adjacency_matrix(
Adjacency_matrix_true,
mode = "directed",
diag = FALSE
)

plot(
  g_true,
  vertex.size = 20,
  vertex.label.cex = 0.8,
  edge.arrow.size = 0.5,
  main = "True DAG Structure"
)
```

With the simulated dataset and prior hyperparameters specified above, we fit the Bayesian LiNGAM model using `BayesDAG().` Each MCMC iteration \[t\] produces:

- Adjacency matrix $E^{(t)}\in \{0,1\}^{p\times p}$
- Causal effect matrix $B^{(t)}\in \mathbb{R}^{p\times p}$
- Mixture parameters for the error model, with component specific means and variances/precisions stored in matrices (e.g., $\mu^{(t)},\tau^{(t)}\in \mathbb{R}^{p\times M}$, where column $k$ corresponds to mixture component $k$).

For posterior summaries, each iteration's matrices are flattened (vectorized) into a single parameter vector. Concretely, we apply $vec(\cdot)$ to stack all entries of $E^{(t)}$, $B^{(t)}$, $\mu^{(t)}$, and $\tau^{(t)}$, and stack these vectors over iterations to form a matrix of samples where the rows are iterations and columns correspond to fixed matrix entries (one column for $B_{ij}$ or $\mu_{ik}$). This makes element wise quantities, edge probabilities, HPD/credible intervals, and other summaries straightforward to compute.

```{r example}
#######################################
# Parameter Initialization
#######################################
B_seed = directlingam_seed(data_matrix)
seed_edge_threshold = 0.05   # drop |seed coef| below this
seed_rho_target = 0.95   # rescale if spectral radius exceeds this
seed_settle_sweeps = 30     # mixture Gibbs sweeps before handing off
#######################################
# Initialization
#######################################  
init_state = init_from_seed(
  B_seed, data_matrix, N, num_covariates, M,
  params$a_mu, params$b_mu, params$a_tao, params$b_tao, params$alpha,
  params$a_gamma, params$b_gamma, params$a_gamma_1, params$b_gamma_1,
  edge_threshold = seed_edge_threshold,
  rho_target     = seed_rho_target,
  settle_sweeps  = seed_settle_sweeps
)
#######################################
# Run Bayesian DAG Sampler
#######################################

results_list = BayesDAG(
      data_matrix,
      params$a_mu, params$b_mu,
      params$a_gamma, params$b_gamma,
      params$a_tao, params$b_tao,
      params$a_og_tao, params$b_og_tao,
      params$a_gamma_1, params$b_gamma_1,
      params$alpha,
      M,
      num_iter,
      init_Adjacency     = init_state$Adjacency_matrix,
      init_Causal_effect = init_state$Causal_effect_matrix,
      init_mu            = init_state$mu_mat,
      init_tao           = init_state$tao_mat,
      init_pi            = init_state$pi_mat,
      init_Z             = init_state$Z_matrix,
      init_gamma_1       = init_state$gamma_1,
      init_gamma_result  = init_state$gamma_result
    )

#######################################
# Extract posterior outputs
#######################################
log_likelihood_list = results_list$log_likelihood_list
Adjacency_matrix_list = results_list$Adjacency_matrix_list 
Causal_effect_matrix_list = results_list$Causal_effect_matrix_list
gamma_list = results_list$gamma_list
gamma_1_list = results_list$gamma_1_list
mu_matrix_list = results_list$mu_matrix_list
tao_matrix_list = results_list$tao_matrix_list
pi_matrix_list = results_list$pi_matrix_list
```

To obtain a representative estimate of the graph structure, we use the function `point_est_graph(),` which selects the posterior weighted medoid, the graph that minimizes the weight distance to all other sampled adjacency matrices. Users may select one of the build in distances or supply their own custom functions. The available distance metrics are:

- **Structural Hamming Distance (SHD):**  
  Counts the number of single edge edits (additions, deletions, or reversals)
  required to transform one graph into another.

- **Structural Intervention Distance (SID):**  
  Measures how many node pairs \((i, j)\) imply different intervention
  distributions P($\mathbf{Y}_j$ | do($\mathbf{Y}_i$)).  
  (Applicable only when the posterior graphs are DAGs)

Users may also specify
```{r}
dist_type = 'custom'
dist_fun = function(A,B){...}
```
where $A$ and $B$ are $p\times p$ adjacency matrices. The function must return a non negative scalar distance. 

```{r}
#############################################
# Best Graph Structure determined through shd
#############################################
Adjacency_matrix_shd = point_est_graph(Adjacency_matrix_list, dist_type = 'shd')
Adjacency_matrix_shd
```

```{r, warning=FALSE, message=FALSE}
############################################
# Best Graph Structure determined through sid
############################################
if (requireNamespace("SID", quietly = TRUE)) {
  Adjacency_matrix_sid = point_est_graph(Adjacency_matrix_list, dist_type = "sid")
  Adjacency_matrix_sid
} else {
  message("SID not installed. Install SID (and possibly Bioconductor graph/RBGL) to run SID.")
}
```
As an example for a custom function, we could count the total number of edge mismatches.

```{r}
custom_edge_mismatch = function(A, B) {
  return(sum(abs(A - B)))
}

Adjacency_matrix_custom = point_est_graph(Adjacency_matrix_list, dist_type = 'custom', dist_fun = custom_edge_mismatch)
Adjacency_matrix_custom
```

We observe that the three structures selected using different distance metrics are all identical.

```{r}
Adjacency_matrix_true
```

In addition to choosing the best possible graphs through distance based medoids, users may evaluate how often a graph appears in structure appears in the posterior samples. This provides posterior support for any candidate DAG, such as the true graph. 

The function `posterior_network_motif()` computes the posterior mass (relative frequency) of a graph by checking whether all its edges are present on each of the posterior graph structures. Using the adjacency samples from the acyclic example above, we can compute the posterior mass assigned to the true graph:

```{r}
true_graph_structure = igraph::graph_from_adjacency_matrix(Adjacency_matrix_true)
posterior_network_motif(true_graph_structure, Adjacency_matrix_list)
```

Since the value is at 0.993, this indicates that the sampler visited the true graph structure repeatedly on every possible posterior graph structure. 

To assess whether the sampler has mixed well and is exploring the parameter space adequately, we inspect the trace plot for the log likelihood. The log likelihood trace is especially useful, as it reflects convergence of the entire parameter configuration rather than a single component. Stable behavior in both the parameter traces and the log likelihood provides strong evidence that the chain has reached stationarity. The trace plot for the log likelihood is plotted below:

```{r}
plot(log_likelihood_list[,1][(0.75*num_iter):num_iter], type='l', xlab = 'Iterations', ylab = 'log likelihood values')
```

As shown above, the log-likelihood remains within a relatively stable range
after burn-in, suggesting that the sampler has reached a stationary regime.
This provides evidence that the retained posterior samples are suitable for
summarizing the remaining model parameters.

Posterior interval estimates are obtained using 
\texttt{posterior\_interval\_est()}, which computes highest posterior density
(HPD) intervals and equal-tailed credible intervals column-wise for each
element of the sampled parameter matrices. To illustrate posterior uncertainty
in the causal-effect coefficients, we examine the nonzero entries of the true
causal-effect matrix $B$.

The figures below display the posterior estimates together with their 95\%
HPD and equal-tailed credible intervals. For each causal edge, the red
$\times$ marks the true value of the corresponding causal effect coefficient,
while the black point and vertical interval summarize the posterior estimate
and uncertainty. This allows each estimated effect to be compared directly
with its ground-truth value. Overall, the posterior estimates lie close to the
true effects, and the intervals generally cover the corresponding ground-truth
values, indicating accurate recovery of the causal-effect magnitudes in this
example.

```{r}
#######################################
# Posterior intervals
#######################################

Causal_effect_matrix_summary =
  posterior_interval_est(Causal_effect_matrix_list, level = 0.95)

hpd_matrix_acyclic = Causal_effect_matrix_summary$hpd_matrix
ci_matrix_acyclic  = Causal_effect_matrix_summary$ci_matrix


#######################################
# True nonzero causal effects
#######################################

# R vectorizes matrices column-by-column
true_vec = as.vector(Causal_effect_matrix_true)

# Indices of all truly nonzero causal effects
true_idx = which(true_vec != 0)

# Convert vector indices back to matrix coordinates
coords = arrayInd(
  true_idx,
  .dim = dim(Causal_effect_matrix_true)
)

# Convention: B_ij is the effect j -> i
true_effects_df = data.frame(
  coef_index = true_idx,
  child  = coords[, 1],
  parent = coords[, 2],
  truth  = true_vec[true_idx]
)

true_effects_df$edge = paste0(
  true_effects_df$parent,
  " -> ",
  true_effects_df$child
)

#true_effects_df


#######################################
# Check dimensions before matching
#######################################

if (ncol(hpd_matrix_acyclic) != length(true_vec)) {
  stop("HPD matrix ordering/dimensions do not match Causal_effect_matrix_true.")
}

if (nrow(ci_matrix_acyclic) != length(true_vec)) {
  stop("CI matrix ordering/dimensions do not match Causal_effect_matrix_true.")
}


#######################################
# HPD intervals for TRUE causal effects
#######################################

hpd_sub = t(
  hpd_matrix_acyclic[, true_idx, drop = FALSE]
)

colnames(hpd_sub) = c("lower", "upper")

data_hpd = data.frame(
  edge  = true_effects_df$edge,
  truth = true_effects_df$truth,
  lower = hpd_sub[, "lower"],
  upper = hpd_sub[, "upper"]
)

# Midpoint of HPD interval
data_hpd$estimate =
  (data_hpd$lower + data_hpd$upper) / 2

# Keep edge order fixed
data_hpd$edge =
  factor(data_hpd$edge, levels = data_hpd$edge)


#######################################
# Plot HPD intervals
#######################################

ggplot(data_hpd, aes(x = edge, y = estimate)) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.2
  ) +
  geom_point(size = 3) +

  # True causal effect
  geom_point(
    aes(y = truth),
    shape = 4,
    size = 4,
    stroke = 1.3,
    color = "red"
  ) +

  labs(
    x = "True Causal Edge",
    y = "Causal Effect",
    title = "Posterior Estimates with 95% HPD Intervals"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )


#######################################
# 95% credible intervals for TRUE effects
#######################################

ci_sub =
  ci_matrix_acyclic[true_idx, , drop = FALSE]

data_ci = data.frame(
  edge     = true_effects_df$edge,
  truth    = true_effects_df$truth,
  lower    = ci_sub[, 1],
  estimate = ci_sub[, 2],
  upper    = ci_sub[, 3]
)

data_ci$edge =
  factor(data_ci$edge, levels = data_ci$edge)


#######################################
# Plot credible intervals
#######################################

ggplot(data_ci, aes(x = edge, y = estimate)) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.2
  ) +
  geom_point(size = 3) +

  # True causal effect
  geom_point(
    aes(y = truth),
    shape = 4,
    size = 4,
    stroke = 1.3,
    color = "red"
  ) +

  labs(
    x = "True Causal Edge",
    y = "Causal Effect",
    title = "Posterior Estimates with 95% Credible Intervals"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )


#######################################
# Coverage
#######################################

data_hpd$covered =
  data_hpd$truth >= data_hpd$lower &
  data_hpd$truth <= data_hpd$upper

data_ci$covered =
  data_ci$truth >= data_ci$lower &
  data_ci$truth <= data_ci$upper

cat(
  "HPD coverage:",
  sum(data_hpd$covered),
  "out of",
  nrow(data_hpd),
  "\n"
)

cat(
  "Credible interval coverage:",
  sum(data_ci$covered),
  "out of",
  nrow(data_ci),
  "\n"
)
```

## Cyclic (DCG) Example

Finally, to illustrate the cyclic Bayesian sampler, we generate a directed
cyclic graph (DCG) containing at least one directed cycle. The graph is
generated under the disjoint-cycle restriction used by `BayesDCG`, so that
no two directed cycles share a node.

For each included edge \(j \to i\), the corresponding nonzero causal-effect
coefficient is generated according to

\[
|B_{ij}| \sim \mathrm{Unif}(0.4,0.9),
\]

with its sign chosen independently to be positive or negative with equal
probability. For excluded edges, \(B_{ij}=0\).

Because feedback is present in a cyclic model, we additionally require the
causal-effect matrix to satisfy the stability condition

\[
\rho(B)
=
\max_{\lambda \in \operatorname{eig}(B)} |\lambda|
< 1.
\]

Only stable causal-effect matrices satisfying this condition are used to
generate the data. This also guarantees that \(I-B\) is invertible.

Given the causal-effect matrix \(B\) and structural error matrix
\(\epsilon\), the observed data are generated from

\[
Y=(I-B)^{-1}\epsilon.
\]

The resulting data therefore arise from a stable linear non-Gaussian cyclic
structural equation model with disjoint directed cycles.

```{r}
#######################################
# Simulation and MCMC settings
#######################################

N = 200
num_covariates = 10
M = 5
num_iter = 80000
burn_in_iterations = 50000


#######################################
# Hyperparameter setup
#######################################

params = list(
  a_mu      = 0,
  b_mu      = 2,
  a_gamma   = 1,
  b_gamma   = 20,
  a_gamma_1 = 0.5,
  b_gamma_1 = 0.5,
  a_tao     = 2,
  b_tao     = 1,
  alpha     = 1
)


#######################################
# Generate DCG Example
#######################################

example_list = generates_examples_DCG(
  num_covariates = num_covariates,
  N              = N,
  M_input        = 2,          # components used to generate the errors
  prob_sparsity  = 0.90,       # target edge density = 0.10
  seed_input     = 21,
  n_cycles       = 2,          # four vertex-disjoint directed cycles
  len_range      = c(2, 4),
  mag_range      = c(0.4, 0.9),
  prob_positive  = 0.5,
  tol_sing       = 0.05,
  rho_max        = 0.95
)


#######################################
# Extract generated data
#######################################

data_matrix = example_list$data_matrix
Adjacency_matrix_true = example_list$Adjacency_matrix_true
Causal_effect_matrix_true = example_list$Causal_effect_matrix_true
Z_matrix_true = example_list$Z_matrix_true

cycles_true = example_list$cycles
rho_true = example_list$rho


#######################################
# Inspect generated DCG
#######################################

cycles_true
rho_true
```


```{r}
#######################################
# Parameter Initialization
#######################################

B_seed = directlingam_seed(data_matrix)

seed_edge_threshold = 0.05   # drop |seed coef| below this
seed_rho_target     = 0.95   # stabilize initialization if needed
seed_settle_sweeps  = 30     # mixture Gibbs sweeps before handing off

#######################################
# Initialization
#######################################

init_state = init_from_seed(
  B_seed, data_matrix, N, num_covariates, M,
  params$a_mu, params$b_mu,
  params$a_tao, params$b_tao,
  params$alpha,
  params$a_gamma, params$b_gamma,
  params$a_gamma_1, params$b_gamma_1,
  edge_threshold = seed_edge_threshold,
  rho_target     = seed_rho_target,
  settle_sweeps  = seed_settle_sweeps
)


#######################################
# Run Bayesian DCG Sampler
#######################################

results_list = BayesDCG(
  data_matrix = data_matrix,

  a_mu      = params$a_mu,
  b_mu      = params$b_mu,
  a_gamma   = params$a_gamma,
  b_gamma   = params$b_gamma,
  a_tao     = params$a_tao,
  b_tao     = params$b_tao,
  a_gamma_1 = params$a_gamma_1,
  b_gamma_1 = params$b_gamma_1,
  alpha     = params$alpha,

  M        = M,
  num_iter = num_iter,

  burn_in_iterations = burn_in_iterations,

  init_Adjacency     = init_state$Adjacency_matrix,
  init_Causal_effect = init_state$Causal_effect_matrix,
  init_mu            = init_state$mu_mat,
  init_tao           = init_state$tao_mat,
  init_pi            = init_state$pi_mat,
  init_Z             = init_state$Z_matrix,
  init_gamma_1       = init_state$gamma_1,
  init_gamma_result  = init_state$gamma_result
)


#######################################
# Extract posterior outputs
#######################################

log_likelihood_list        = results_list$log_likelihood_list
Adjacency_matrix_list      = results_list$Adjacency_matrix_list
Causal_effect_matrix_list  = results_list$Causal_effect_matrix_list
gamma_list                 = results_list$gamma_list
gamma_1_list               = results_list$gamma_1_list
mu_matrix_list             = results_list$mu_matrix_list
tao_matrix_list            = results_list$tao_matrix_list
pi_matrix_list             = results_list$pi_matrix_list

# Number of retained posterior graph samples
posterior_sample_length = nrow(Adjacency_matrix_list)
```

Before examining posterior summaries, again we will visualize the true underlying DCG used in the simulation. The plot below displays the ground truth adjacency structure.

```{r}
#######################################
# Build directed graph from adjacency matrix
#######################################

g_true = graph_from_adjacency_matrix(
Adjacency_matrix_true,
mode = "directed",
diag = FALSE
)

plot(
  g_true,
  vertex.size = 20,
  vertex.label.cex = 0.8,
  edge.arrow.size = 0.5,
  main = "True DCG Structure"
)
```

With the simulated dataset and prior hyperparameters specified above, we now fit the Bayesian LiNGAM model using `BayesDCG().` The function, similar to `BayesDAG(),` returns same posterior samples for 

- Adjacency matrices,
- Causal effect matrices, 
- Mixture parameters for the error model, with component specific means and variances/precisions stored in matrices. 


To obtain a representative estimate of the graph structure, we again use the function `point_est_graph(),` which selects the posterior weighted medoid under a chosen distance metric. We use the same distance metrics, but the Structural Intervention Distance (SID) is only applicable when the sampled graphs are DAGs. Users may select shd or supply their own custom functions. 

```{r,warning=FALSE, message=FALSE, error = TRUE}
# SID is shown for completeness, it applies only when all posterior graphs are DAGs. If any sampled graph contains a cycle, SID-based selection will produce an error.
Adjacency_matrix_sid = point_est_graph(Adjacency_matrix_list, dist_type = 'sid')
```

```{r}
Adjacency_matrix_shd = point_est_graph(Adjacency_matrix_list,dist_type = 'shd')
Adjacency_matrix_shd
```

Using the same custom function as the acyclic case, we get following graph structure. 

```{r}
Adjacency_matrix_custom = point_est_graph(Adjacency_matrix_list, dist_type = 'custom', dist_fun = custom_edge_mismatch)
Adjacency_matrix_custom
```

In this example, the selected graphs coincide with the true adjacency matrix:

```{r}
Adjacency_matrix_true
```

In addition to choosing the best possible graphs through distance based medoids, users similarly may evaluate how often a graph appears in structure appears in the posterior samples using `posterior_network_motif().` This provides posterior support for any candidate DCG, such as the true graph. 

```{r}
true_graph_structure = igraph::graph_from_adjacency_matrix(Adjacency_matrix_true)
posterior_network_motif(true_graph_structure, Adjacency_matrix_list)
```

Since the value is at 1, this indicates that the sampler visited the true graph structure repeatedly on every possible posterior graph structure.

Similar to the acyclic case, to check whether the sampler has mixed well and is exploring the parameter space well, we again inspect the log likelihood. Below we plot the log likelihood for each iteration to assess mixing and posterior behavior:

```{r}
plot(log_likelihood_list[,1][(0.75*num_iter):num_iter], type='l', xlab = 'Iterations', ylab = 'log likelihood values')
```
As seen overall, the log_likelihoods stay in the same general area indicating the sampler has reached a stationary regime, indicating posterior summaries for the other parameters are from a well converged chain. Similar to the acyclic case, we can summarize uncertainty in the causal effect coefficients using `posterior_interval_est(),` which computes HPD and equal tailed credible intervals for each specific parameter entry represented as columns of the matrix outputs. 

```{r}
# Compute HPD and CI summaries for causal effect matrix
Causal_effect_matrix_summary = posterior_interval_est(Causal_effect_matrix_list, level = 0.95)
hpd_matrix_cyclic = Causal_effect_matrix_summary$hpd_matrix
ci_matrix_cyclic = Causal_effect_matrix_summary$ci_matrix

#######################################
# Extracting nonzero HPD intervals
#######################################

par(mfrow=c(2,1))

nonzero_cols = which(colSums(hpd_matrix_cyclic) != 0)

hpd_sub = t(hpd_matrix_cyclic[, nonzero_cols, drop = FALSE])
colnames(hpd_sub) = c("lower", "upper")

# build data frame
data_cyclic = as.data.frame(hpd_sub)
data_cyclic$x = factor(seq_len(nrow(data_cyclic)))  # index for plotting
data_cyclic$mid = (data_cyclic$lower + data_cyclic$upper) / 2

###############################################
# HPD interval plot (cyclic)
###############################################
ggplot(data_cyclic, aes(x = x, y = mid)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_hline(yintercept = 0.7143305, linetype = "dashed", color = "red") +
  labs(
    x = "Nonzero causal-effect coefficient (index)",
    y = "Causal Weight Estimate with HPD Interval"
  ) +
  theme_minimal()

#######################################
# Credible intervals
#######################################

data_2 = data.frame(ci_matrix_cyclic[which(rowSums(ci_matrix_cyclic)!=0),])
x = 1:nrow(data_2)
data_2 = cbind(x,data_2)

ggplot(data_2, aes(x = x, y = X2)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0.7143305, linetype = "dashed", color = "red") +
  geom_errorbar(aes(ymin = X1, ymax = X3), width = 0.2) +  
  labs(y = "Causal Weight Estimate with 95% CI", x = "Nonzero causal-effect coefficient (index)") +
  theme_minimal()

```


## Documentation (vignette + appendix)

For a complete tutorial **and the supplemental material** (appendix with additional simulations and implementation details), see the package vignette:
```{r}
#browseVignettes("cyclinbayes")
```


























